IML Term project

### 2022-12-02 main_markdown.rmd
LDA- ja QDA-testaus.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

@@

# Luodaan class2
npf$class2 <- factor("event", levels = c("nonevent", "event"))
npf$class2[npf$class4 == "nonevent"] <- "nonevent"

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
```

@@

# Luodaan class2, 0 = nonevent, 1 = event
npf$class2 <- ifelse(npf$class4 == "nonevent", 0, 1)

@@

# Puolitetaan training data jotta saadaan realistinen testausdata
set.seed(42)
idx <- sample(nrow(npf), nrow(npf)/2)
npf_train <- npf[idx, ]
npf_test <- npf[-idx, ]
```

```{r functions}

# mean squared error
mse <- function(yhat, y) sqrt(mean((y - yhat)**2))

## accuracy
accuracy <- function(pred, y) mean(ifelse(pred >= 0.5, 1, 0) == y)

## perplexity. Pakotetaan todennäköisyydet tässä välille [0.995, 0.005]
perplexity <- function(pred, y){
  pred <- ifelse(pred != 0 & pred != 1, pred, ifelse(pred == 0, 0.005, 0.995 ))
  exp(-mean(log(ifelse(y == 1, pred, 1 - pred)))) }

# cross-validate
# split n items into k folds of roughly equal size
kpart <- function(n, k) {
    rep_len(1:k, length.out = n)
}

# Find cross-validation predictions
cv <- function(
               formula, # Formula specifying which variables to use
               data, # Dataset
               model = lm, # Type of model to train (as a function)
               n = nrow(data), # number of rows in the data matrix
               k = min(n, 10), # number of cross-validation folds
               split = kpart(n, k), # the split of n data items into k folds
               ## function to train a model on data
               train = function(data) model(formula, data = data),
               ## function to make predictions on the trained model
               pred = function(model, data) predict(model, newdata = data, type = "raw")[,2]) {
    yhat <- NULL
    for (i in 1:k) {
        ## go through all folds, train on other folds, and make a prediction
        mod <- train(data[split != i, ])
        if (is.null(yhat)) {
            ## initialise yhat to something of correct data type,
            yhat <- pred(mod, data)
        } else {
            yhat[split == i] <- pred(mod, data[split == i, ])
        }
    }
    yhat # finally, output cross-validation predictions
}

# loocv <- function(
#                formula, # Formula specifying which variables to use
#                data, # Dataset
#                model = lm, # Type of model to train (as a function)
#                ## function to train a model on data
#                train = function(data) model(formula, data = data),
#                ## function to make predictions on the trained model
#                pred = function(model, data) predict(model, newdata = data, type = "raw")[,2]) {
#     yhat <- NULL
#     for (i in 1:nrow(data)) {
#         ## go through all stations, train on other stations, and make a prediction
#         mod <- train(data[-i, ])
#         if (is.null(yhat)) {
#             ## initialise yhat to something of correct data type,
#             yhat <- pred(mod, data)
#         } else {
#             yhat[i] <- pred(mod, data[i, ])
#         }
#     }
#     yhat # finally, output cross-validation predictions
# }

# Classifiers

## Testing classifiers as they are

### Naive Bayes

```{r naive bayes}

library(e1071)

nb.fit <- naiveBayes(class2 ~ . , data = npf_train, laplace = 1)
nb.pred.cv <- cv(class2 ~ ., npf, naiveBayes) # Hidas prosessi, tehdään siksi vaan kerran

nb_acc <- data.frame(train = accuracy(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                     test = accuracy(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                     CV = accuracy(nb.pred.cv, npf$class2))

# Kaikkiin tulee Inf
nb_perp <- data.frame(train = perplexity(predict(nb.fit, newdata = npf_train, type = "raw")[,2], npf_train$class2),
                      test = perplexity(predict(nb.fit, newdata = npf_test, type = "raw")[,2], npf_test$class2),
                      CV = perplexity(nb.pred.cv, npf$class2))

kable(cbind(nb_acc,
            nb_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))

### LDA
```{r lda}

library(MASS)

lda.fit <- lda(class2 ~ . , data = npf_train)

# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
lda.pred.cv <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- lda(class2 ~ ., npf[split != i, ])
  lda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}

lda_acc <- data.frame(train = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(lda.pred.cv, npf$class2))

lda_perp <- data.frame(train = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(lda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(lda.pred.cv, npf$class2))

kable(cbind(lda_acc,
            lda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))

### QDA
```{r qda}

library(MASS)

qda.fit <- qda(class2 ~ . , data = npf_train)

# Cross-validation parametrit
n = nrow(npf) # number of rows in the data matrix
k = min(n, 10) # number of cross-validation folds
split = kpart(n, k) # the split of n data items into k folds
qda.pred.cv <- rep(0,n)

# cv
for (i in 1:k) {
  ## go through all folds, train on other folds, and make a prediction
  mod <- qda(class2 ~ ., npf[split != i, ])
  qda.pred.cv[split == i] <- predict(mod, npf[split == i, ])$posterior[,2]
}

qda_acc <- data.frame(train = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = accuracy(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = accuracy(qda.pred.cv, npf$class2))

# Taas Inf :(
qda_perp <- data.frame(train = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_train$class2),
                      test = perplexity(predict(qda.fit, newdata = npf_train)$posterior[,2], npf_test$class2),
                      CV = perplexity(qda.pred.cv, npf$class2))

kable(cbind(qda_acc,
            qda_perp)) %>%
  add_header_above(c("Accuracy" = 3, "Perplexity" = 3))
